Our Title

### 2015-05-04 draft.tex
Added in James K's section on basic nuclear physics.

@@

\subsection{??????   Assign to someone  ????????}

@@

\subsection{???????   Need discussion of criticality  ??????}

A fission device operates on the principle that fissile materials (list of materials) will undergo fission when stuck by neutrons of sufficient kinetic energy. A nuclide which undergoes fission will release between 100-200 MeV in addition to 2-3 neutrons, which may be utilized to cause subsequent fissions in other fissile nuclei \cite{krane1987introductory}. The minimum amount of fissile material needed in order to sustain a chain reaction is known as the critical mass, which depends on the type, density, and shape of material used. A tamper, which is a high Z material which acts as a neutron reflector, is typically used in order to reduce the critical mass.

\subsection{??????? Need discussion of radioactive decay   ?????}

@@

In beta decay, the electron's kinetic energy varies, with an average of 1/3 of the maximum electron energy, while the remaining energy is carried off by the nearly undetectable electron antineutrino.

@@

The minimum amount of special nuclear material needed to produce a fission weapon is called the critical mass. This value can depend on many different features including isotope, geometry of the material, micro-structure of the material, and device design. Given the many variables in play it would be infeasible to include the critical mass of every possible weapons configuration here. However, to give a general idea, \autoref{tab:critical_mass} presents the critical mass of two different designs for several types of weapons. Additionally the yields of fission devices can vary based on many factors including, type of material, device design, amount, efficiency, and many other factors. Yields are typically on the order of kilotons \cite{Moody2014}.

@@

\textbf{Electron Volt:} A unit of energy equivalent to the work required to accelerate an electron through a potential difference of 1 volt.

>>>>

This section is meant to serve as a basic guide to technical concepts necessary for understanding nuclear weapons. Although the content has been distilled to the most fundamental information and is deliberately simplified, it is critical to understanding the scientific context that nuclear weapons find themselves in \cite{Prussin2014}.

@@

\subsection{??????? Isotope and nuclear structure section  ????}

@@

Nuclear fission is the process in which a nucleus splits into two nuclei of roughly equal mass \cite{krane1987introductory,Lamarsh2013}. This is the dominant reaction in nuclear weapons that produces the incredible amount of energy released upon detonation. The two isotopes used in fission-based systems as fuel are \ce{^{235}U} and \ce{^{239}Pu}. These fuel isotopes are used individually in a single weapon, not as a mixture. In weapons, neutrons are used to begin the fission process in nuclei, which when they fission also release excess neutrons, which go on to spark further fission reactions. This is known as the fission chain reaction. Controlling this chain reaction is the business of nuclear reactors, while accelerating it as much as possible is the purpose of a nuclear weapon \cite{Duderstadt1976}.

It is also important to understand that the two fuel isotopes are not naturally found in forms that can be directly used to fuel a nuclear weapon \cite{Benedict1981,Lewis2008}. Instead, \ce{^{235}U} is found in extremely small amounts in natural uranium deposits and must be further enriched and concentrated for use in a weapon. \ce{^{239}Pu}, on the other hand, is not naturally occurring and must be produced in a nuclear reactor specially designed for this purpose. Because of this, \ce{^{235}U} is the easier to acquire and use of the two fuel isotopes, and thus is of more concern.

A fission reactor or device operates on the principle that fissile materials (list of materials) will undergo fission when stuck by neutrons of sufficient kinetic energy. A nuclide which undergoes fission will release between 100-200 MeV in addition to 2-3 neutrons, which may be utilized to cause subsequent fissions in other fissile nuclei \cite{krane1987introductory}. The minimum amount of fissile material needed in order to sustain a chain reaction is known as the critical mass, which depends on the type, density, and shape of material used. A tamper, which is a high Z material which acts as a neutron reflector, is typically used in order to reduce the critical mass.

\subsection{Criticality}

Criticality in a system is defined as the state in which the rate of reactions in the system is constant, producing a constant energy output over time. In a fission system, criticality is a constant rate of fission reactions over time, and is also known as a controlled fission reaction \cite{Duderstadt1976}. This is the dominant operating state of nuclear reactors that maintain a constant power level during the majority of their operational lives \cite{Ott1983}. A supercritical system is a system where the reaction rate is constantly increasing, and a subcritical system is thus a system where the reaction rate is constantly decreasing.

Clearly, it is important to understand that nuclear weapons rely upon the fact that their nuclear fuel can be arranged in such a way that it becomes violently supercritical when the user specifies, and not before \cite{Lewis2008}. Therefore, in terms of criticality, it can be said that the nuclear core of a weapon is initially subcritical before detonation, and  is rendered highly supercritical at the exact moment of detonation. This is fundamentally what allows the weapon to release such earth-shaking amounts of energy nearly instantly when it is used, but to be stored safely before then \cite{Prussin2014,Defense1998}.

The minimum amount of special nuclear material needed to produce the critical core of a  fission weapon is called the critical mass. This value can depend on many different features including isotope, geometry of the material, micro-structure of the material, and device design. Given the many variables in play it would be infeasible to include the critical mass of every possible weapons configuration here. However, to give a general idea, \autoref{tab:critical_mass} presents the critical mass of two different designs for several types of weapons. Additionally the yields of fission devices can vary based on many factors including, type of material, device design, amount, efficiency, and many other factors. Yields are typically on the order of kilotons \cite{Moody2014}.

Fusion is fundamentally the inverse process of fission. Fusion is the process wherein two light nuclei combine to form a single, heavier nucleus, and in doing so release a large amount of energy, possibly in addition to neutrons or various charged particles \cite{krane1987introductory,Dolan1982}. In thermonuclear weapons, fusion is typically used to \enquote{boost} the total energy production of a weapon significantly beyond the release otherwise achieved through fission alone. Fundamentally, the fusion reaction provides a large initial supply of neutrons that allows the fission process to begin at a much higher rate than it could alone \cite{Prussin2014}.

\subsection{Energy Release}

Typical nuclear reactions release energies on the order of 1-10 MeV \cite{Prussin2014}. In comparison, typical chemical reactions release energy on the order of a single eV per molecule, and perfectly combusted TNT, a common high-explosive, releases approximately 35 eV per molecule \cite{Prussin2014}. Clearly the amount of energy released in nuclear reactions is many orders of magnitude greater than typical chemical reactions more familiar to society. However, the energy released in fission, the primary nuclear reaction in nuclear weapons, is 100 times that of typical nuclear reactions (approximately 180 MeV per reaction) \cite{krane1987introductory,Loveland2005}. Combining the incredible energy released per fission reaction with the density of the fuel in nuclear weapons, it becomes readily apparent that nuclear weapons contain an energy source greater than any other on earth. Combined  with the incredibly short timescales on which nuclear reactions occur (on the order of 10\(^{-7}\) s - 10\(^{-6}\) s),  nuclear weapons become tools of unimaginable power and destruction \cite{Loveland2005,Cochran1994}.

\subsection{Radioactive Decay}

For all elements, there are isotopes which are not naturally occurring. The reason that these isotopes are not observed in nature is that they are radioactive. A radioactive isotope, due to is structure, is energetically unstable, to the point that it will emit excess nucleons or energy to become stable \cite{Loveland2005}. This process of reaching a more stable nuclear structure is known as radioactive decay. The three most common modes of radioactive decay are referred to as alpha (\(\alpha\)) decay, beta (\(\beta\)) decay, and gamma (\(\gamma\)) decay \cite{Loveland2005}. In general, isotopes that are very energetically unstable decay faster and with greater releases of energy, whereas isotopes that are not far from stability decay more slowly and with lower-energy releases. This makes very unstable isotopes very dangerous; less unstable isotopes are less of a radiation hazard \cite{krane1987introductory}.

Alpha and beta decay change the structure of the nucleus of the atom, by emitting or transforming excess nucleons to stabilize the nucleus. The change in the nucleus' structure literally changes it into a different element altogether \cite{krane1987introductory}. Alpha decay allows heavy (A \textgreater \(\approx\) 110) nuclei to stabilize themselves by shedding excess nuclei. In alpha decay, the nucleus emits an \(\alpha\) particle, also known as the nucleus of a \ce{^{4}He} atom. This particle is composed of two protons and two neurons, and typically has a kinetic energy of 4 - 10 MeV \cite{krane1987introductory}.

Beta decay allows nuclei with an excess of either protons or neutrons to stabilize themselves by transforming an excess proton into a neutron, or vice-versa, releasing a neutrino in the process \cite{krane1987introductory}. In beta decay,  the nucleus emits a \(\beta\) particle, which is either an electron, or its positively-charged counterpart, the positron. \(\beta\) particles typically have a kinetic energy of several hundred keV;   the electron's kinetic energy varies, with an average of 1/3 of the maximum electron energy, while the remaining energy is carried off by a nearly undetectable electron antineutrino \cite{krane1987introductory}.

Gamma decay allows nuclei with a high-energy internal configuration of their nucleons to stabilize by shifting that configuration to one lower in energy. The extra energy is released in the form of a high-energy electromagnetic wave, known as a gamma ray; the identity of the isotope is unchanged. Gamma rays have a much larger range of typical energies than alpha or beta particles, ranging approximately from 10 keV - 15 MeV  \cite{krane1987introductory}.

@@

@@

Missing...

@@

\textbf{Electron Volt (eV):} A unit of energy equivalent to the work required to accelerate an electron through a potential difference of 1 volt.
